An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-01 poly-embed-journal/poly-embed-journal.tex
Notes pre meeting

\begin{definition}[Embedding a loss]\label{def:loss-embed}
  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ over a representative set $\Sc$ if there exists some injective embedding $\varphi:\Sc\to\reals^d$ such that
  (i) for all $r\in\Sc$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
  % \begin{equation}\label{eq:embed-loss}
  %   r \in \argmin_{r'\in\R} \inprod{p}{\ell(r')} \iff \varphi(r) \in \argmin_{u\in\reals^d} \inprod{p}{L(u)}~.
  % \end{equation}
  \begin{equation}\label{eq:embed-loss}
    r \in \prop{\ell}(p) \iff \varphi(r) \in \prop{L}(p)~.
  \end{equation}
  We simply say $L$ \emph{embeds} $\ell$ if there is some representative set $\Sc$ for $\ell$ such that $L$ embeds $\ell$ over $\Sc$.
\end{definition}

@@

\jessie{Adding new statements}
\begin{definition}[$\Sc(L)$]\label{cons:rep-set}
	Suppose a generic loss $L : \R \to \reals^\Y$ is given with polyhedral negative risk $-\risk{L}$.
	Construct the set $\Sc(L)$ as follows:
	Each facet $F$ of the epigraph of negative risk $E_L$ must be full dimensional by definition of a facet.
	Take $p \in \inter{F}$, and observe that it must by uniquely supported by some hyperplane $p \mapsto (p, L(u))$ for some $u \in \R$.
	Add $u$ to $\Sc(L)$, and repeat, iterating over each of $E_L$'s finitely many facets.
\end{definition}

\begin{proposition}\label{prop:SL-minimal}
	If loss $L: \R \to \reals^\Y_+$ has polyhedral risk, $\Sc(L)$ is a minimal representative set for $L$.
\end{proposition}
\begin{proof}\jessiet{I don't like this proof, but meh.}
	First, we have to show that $\Sc(L)$ is a representative set.
	As facets are closed and the hyperplane $p_F \mapsto (p_F, \inprod{p_F}{L(u_F)})$ supports $E_L$ (epigraph of negative risk) over the entire facet (and therefore $\Gamma(p) \cap \Sc(L) \ni u_F$ for all $p$ such that the hyperplane supports $E_L$ at $p$), we have $\Sc(L)$ representative for $L$.

	If $\Sc(L)$ was not a minimal set, then we could remove some $s$ from $\Sc(L)$ such that $\Sc' := \Sc(L) \setminus \{s\}$ is also a representative set for $L$.
	By the construction of $\Sc(L)$ over facets of $E_L$, we have some distribution $p$ such that the hyperplane $p \mapsto (p, L(s))$ uniquely supports $E_L$.
	We particularly have $\Gamma(p) \cap \Sc' = \emptyset$, since any other element $s'$ of $\Sc'$ being in $\Gamma(p)$ would imply $p \mapsto (p, L(s'))$ also supports $E_L$, but this hyperplane supporting $E_L$ must be unique, so these are equal.
	By construction of $\Sc$, we must have only added at most one of these to the set.
	Therefore, $\Sc(L)$ is a minimal representative set.
\end{proof}

@@

\begin{theorem}[Aurenhammer~\cite{aurenhammer1987power}]\label{thm:aurenhammer}
	A cell complex is affinely equivalent to a convex polyhedron if and only if it is a power diagram.
\end{theorem}

@@

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  \raft{4/2/2021 first part of the proof should be done now.}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$.
  Since $\Sc$ is representative for $\ell$, for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, so $\U$ is representative for $L$.
  \raft{Commented out: the symmetric argument that representation goes both ways under an embedding.  I'm thinking we should expand on this perspective in section 5, maybe introduce a notion of equivalence between any two losses (discrete or otherwise), and define trim in that context.}
  % Taking $\U := \varphi(\Sc)$, we have that $\ell$ if and only if $\U$ is representative of $L$, as $\gamma(p) \cap \Sc \neq \emptyset \iff \exists r \in \Sc$ such that $r \in \gamma(p) \iff \exists r$ such that $\varphi(r) \in \Gamma(p) \cap \U \iff \Gamma(p) \cap \U \neq \emptyset$.
%  To see that $\gamma'(p) \neq \emptyset$ for all $p\in\simplex$, note that by the definition of $\gamma$ as the property elicited by $\ell$ we have some $r \in \gamma(p)$, and by the embedding condition~\eqref{eq:embed-loss}, $\varphi(r) \in \Gamma(p)$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

@@

We match optimality of embedded reports by construction, since the optimal reports support the epigraph of negative Bayes risk on exactly the same distributions, meaning a report and its embedding are optimal at exactly the same distributions.
To verify matching losses, observe $\risk \ell (\delta_y) = \risk L (\delta_y)$ for each $y \in \Y$, and these risks are attained by some report $r$ and its embedding $u_r$, respectively.
\jessie{This feels like it's missing a detail about how $\Sc$ is minimal and how level sets are full-dimensional.}
\end{proof}

@@

\raft{I wonder if we should demote this to a corollary.  We can write the proof of the reverse implication in Proposition~\ref{prop:embed-bayes-risks} like this: first show that if $\risk{L}$ is polyhedral, there is a finite $\U$ which is representative for $L$.  Then show that if $\risk{\ell} = \risk{L}$ then there is some $\Sc$ that is bijective to the particular $\U$ we constructed.  I'm not sure how this will go, but perhaps worth a try, since it would avoid the proof duplication too!}
\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss $\ell : \R \to \reals^\Y_+$ if and only if $\risk{L}$ is polyhedral.
\end{proposition}
\begin{proof}
  If $L$ embeds a discrete $\ell$, Proposition~\ref{prop:embed-bayes-risks} gives us $\risk{L} = \risk{\ell}$ and polyhedral.
  For the converse, let $\risk{L}$ be polyhedral. %; we again examine the proof of Proposition~\ref{prop:embed-bayes-risks}.
  \proposedadd{Form the finite set $\Sc := \Sc(L)$ as in Definition~\ref{cons:rep-set}.
  %For each facet of $E_L$, which is full-dimensional, there must be some interior distribution $p$ such that $p \mapsto (p, \inprod{p}{L(u)})$ uniquely supports $E_L$, and such a $u$ must exist or else the Bayes risk is not attained.  Take $\Sc$ to be the collection of such reports $u$, and observe $\Sc$ is representative for $L$.  
  By Corollary~\ref{cor:representative-embeds-restriction}, $L$ embeds $L|_\Sc$, which is finite by construction.}
%  The projection of the epigraph of $\risk{L}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer} with finitely many cells $C_1,\ldots,C_k$, which we can index by $\R := \{1,\ldots,k\}$.
%  Defining the property $\gamma:\simplex\toto\R$ by $\gamma_r = C_r$ for $r\in\R$, we see that the same construction gives us points $u_r \in\reals^d$ such that $u_r \in \Gamma(p) \iff r \in \gamma(p)$.
%  Define $\ell:\R\to\reals^\Y_+$ by $\ell(r) = L(u_r)$, the same proof shows that $L$ embeds $\ell$.
\end{proof}

@@

\citet{lapin2015top, lapin2016loss, lapin2018analysis} provide the following convex surrogate loss for this problem, which \citet{yang2018consistency} show to be inconsistent:%
\footnote{Yang and Koyejo also introduce a consistent surrogate, but it is non-convex.}
\begin{align}\label{eq:L-2-surrogate}
L^k(u)_y := \left( 1-u_y + \tfrac{1}{k} \textstyle\sum_{i=1}^k (u - e_y)_{[i]} \right)_+~,
\end{align}
where $e_y$ is $1$ in component $y$ and 0 elsewhere.
With our framework, we can say more.
Specifically, while $(L^k,\psi)$ is not consistent for $\elltopk$, since $L^k$ is polyhedral (Lemma~\ref{lem:top-k-polyhedral}), we know from Theorem~\ref{thm:poly-embeds-discrete} that it embeds \emph{some} discrete loss $\ell^k$, and from Theorem~\ref{thm:eps-thick-calibrated} there is a link $\psi'$ such that $(L^k,\psi')$ is calibrated (and consistent) for $\ell^k$.
We therefore turn to deriving this discrete loss $\ell^k$.

>>>>

\begin{definition}[Embedding a loss]\label{def:loss-embed}
  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ over a representative set $\Sc$ if there exists some injective embedding $\varphi:\Sc\to\reals^d$ such that
  (i) for all $r\in\Sc$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\Sc$ we have
  % \begin{equation}\label{eq:embed-loss}
  %   r \in \argmin_{r'\in\R} \inprod{p}{\ell(r')} \iff \varphi(r) \in \argmin_{u\in\reals^d} \inprod{p}{L(u)}~.
  % \end{equation}
  \begin{equation}\label{eq:embed-loss}
    r \in \prop{\ell}(p) \iff \varphi(r) \in \prop{L}(p)~.
  \end{equation}
  We simply say $L$ \emph{embeds} $\ell$ if there is some representative set $\Sc$ for $\ell$ such that $L$ embeds $\ell$ over $\Sc$.
\end{definition}

@@

\jessie{Adding new statements}
jessiet{Make Construction- not definition}
\begin{definition}[$\Sc(L)$]\label{cons:rep-set}
	Suppose a generic loss $L : \R \to \reals^\Y$ is given with polyhedral negative risk $-\risk{L}$.
	Construct the set $\Sc(L)$ as follows:
	Each facet $F$ of the epigraph of negative risk $E_L$ must be full dimensional by definition of a facet.
	Take $p \in \inter{F}$, and observe that it must by uniquely supported by some hyperplane $p \mapsto (p, L(u))$ for some $u \in \R$.
	Add $u$ to $\Sc(L)$, and repeat, iterating over each of $E_L$'s finitely many facets.
\end{definition}

\begin{proposition}\label{prop:SL-minimal}
	If loss $L: \R \to \reals^\Y_+$ has polyhedral risk, $\Sc(L)$ is a minimum representative set for $L$.
\end{proposition}
\begin{proof}
	First, we have to show that $\Sc(L)$ is a representative set.
	As facets are closed, for each facet $F$ and $p_F \in F$, there is a report $u_F \in \Sc(L)$ by construction and closedness of facets so that the hyperplane $p_F \mapsto (p_F, \inprod{p_F}{L(u_F)})$ supports $E_L$ (epigraph of negative risk) exactly on facet (and therefore $u_F \in \Gamma(p) \cap \Sc(L)$ for all $p$ such that the hyperplane supports $E_L$ at $p$), we have $\Sc(L)$ representative for $L$.
	Representativeness follows the facets of $E_L$ union to the simplex.

	If $\Sc(L)$ was not a minimum set, then we could remove some $s$ from $\Sc(L)$ such that $\Sc' := \Sc(L) \setminus \{s\}$ is also a representative set for $L$.
	By the construction of $\Sc(L)$ over facets of $E_L$, this report $s$ corresponds to a facet $F$ of $E_L$ since facets are full-dimensional and thus have well-defined nonempty interiors.
	There is then some distribution $p_F \in \inter F$ such that the hyperplane $p \mapsto (p, L(s))$ uniquely supports $E_L$ at $p_F$.
	We particularly have $\Gamma(p) \cap \Sc' = \emptyset$, since any other element $s'$ of $\Sc'$ being in $\Gamma(p)$ would imply $p \mapsto (p, L(s'))$ also supports $E_L$, but this hyperplane supporting $E_L$ must be unique, so these are equal.
	By construction of $\Sc$, we must have only added at most one of these to the set.
	Therefore, $\Sc(L)$ is a minimum representative set.
\end{proof}

@@

\jessiet{Move}
\begin{theorem}[Aurenhammer~\cite{aurenhammer1987power}]\label{thm:aurenhammer}
	A cell complex is affinely equivalent to a convex polyhedron if and only if it is a power diagram.
\end{theorem}

@@

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  \raft{4/2/2021 first part of the proof should be done now.}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$; take $\U := \varphi(\Sc)$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$.
  Since $\Sc$ is representative for $\ell$, for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, so $\U$ is representative for $L$.
  \raft{Commented out: the symmetric argument that representation goes both ways under an embedding.  I'm thinking we should expand on this perspective in section 5, maybe introduce a notion of equivalence between any two losses (discrete or otherwise), and define trim in that context.}
  % Taking $\U := \varphi(\Sc)$, we have that $\ell$ if and only if $\U$ is representative of $L$, as $\gamma(p) \cap \Sc \neq \emptyset \iff \exists r \in \Sc$ such that $r \in \gamma(p) \iff \exists r$ such that $\varphi(r) \in \Gamma(p) \cap \U \iff \Gamma(p) \cap \U \neq \emptyset$.
%  To see that $\gamma'(p) \neq \emptyset$ for all $p\in\simplex$, note that by the definition of $\gamma$ as the property elicited by $\ell$ we have some $r \in \gamma(p)$, and by the embedding condition~\eqref{eq:embed-loss}, $\varphi(r) \in \Gamma(p)$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

@@

We match optimality of embedded reports by construction, since the optimal reports support the epigraph of negative Bayes risk on exactly the same distributions, meaning a report and its embedding are optimal at exactly the same distributions.
To verify matching losses, consider that the hyperplanes $p \mapsto (p, \inprod{p}{\ell(r)})$ and $p \mapsto (p, \inprod{p}{L(\varphi(r))})$ are exactly equal since they are linear functions matching on a set of positive measure relative to the simplex, and must be equal everywhere.
This implies $\ell(r) = L(\varphi(r))$ for all $r \in \Sc$ such that the hyperplanes support $E_\ell$ and $E_L$ on a facet.
Since $\Sc$ and $\U$ are minimum representative sets, this holds for all $r \in \Sc$.
%To verify matching losses, observe $\risk \ell (\delta_y) = \risk L (\delta_y)$ for each $y \in \Y$, and these risks are attained by some report $r$ and its embedding $u_r$, respectively.
%\jessie{This feels like it's missing a detail about how $\Sc$ is minimal and how level sets are full-dimensional.}
\end{proof}

@@

\raft{I wonder if we should demote this to a corollary.  We can write the proof of the reverse implication in Proposition~\ref{prop:embed-bayes-risks} like this: first show that if $\risk{L}$ is polyhedral, there is a finite $\U$ which is representative for $L$.  Then show that if $\risk{\ell} = \risk{L}$ then there is some $\Sc$ that is bijective to the particular $\U$ we constructed.  I'm not sure how this will go, but perhaps worth a try, since it would avoid the proof duplication too!}
\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss $\ell : \R \to \reals^\Y_+$ if and only if $\risk{L}$ is polyhedral.
\end{proposition}
\begin{proof}
  If $L$ embeds a discrete $\ell$, Proposition~\ref{prop:embed-bayes-risks} gives us $\risk{L} = \risk{\ell}$ and polyhedral.
  For the converse, let $\risk{L}$ be polyhedral. %; we again examine the proof of Proposition~\ref{prop:embed-bayes-risks}.
  Form the finite set $\Sc := \Sc(L)$ as in Definition~\ref{cons:rep-set}.
  %For each facet of $E_L$, which is full-dimensional, there must be some interior distribution $p$ such that $p \mapsto (p, \inprod{p}{L(u)})$ uniquely supports $E_L$, and such a $u$ must exist or else the Bayes risk is not attained.  Take $\Sc$ to be the collection of such reports $u$, and observe $\Sc$ is representative for $L$.  
  By Corollary~\ref{cor:representative-embeds-restriction}, $L$ embeds $L|_\Sc$, which is finite by construction.
%  The projection of the epigraph of $\risk{L}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer} with finitely many cells $C_1,\ldots,C_k$, which we can index by $\R := \{1,\ldots,k\}$.
%  Defining the property $\gamma:\simplex\toto\R$ by $\gamma_r = C_r$ for $r\in\R$, we see that the same construction gives us points $u_r \in\reals^d$ such that $u_r \in \Gamma(p) \iff r \in \gamma(p)$.
%  Define $\ell:\R\to\reals^\Y_+$ by $\ell(r) = L(u_r)$, the same proof shows that $L$ embeds $\ell$.
\end{proof}

@@

\citet{lapin2015top, lapin2016loss, lapin2018analysis} provide the following convex surrogate loss for this problem, which \citet{yang2018consistency} show to be inconsistent:%
\footnote{\citet{yang2018consistency} also introduce a consistent surrogate, but it is non-convex.}
\begin{align}\label{eq:L-2-surrogate}
L^k(u)_y := \left( 1-u_y + \tfrac{1}{k} \textstyle\sum_{i=1}^k (u - e_y)_{[i]} \right)_+~,
\end{align}
where $e_y$ is $1$ in component $y$ and 0 elsewhere.
With our framework, we can say more.
Specifically, while $(L^k,\psi)$ is not consistent for $\elltopk$, since $L^k$ is polyhedral (Lemma~\ref{lem:top-k-polyhedral}), we know from Theorem~\ref{thm:poly-embeds-discrete} that it embeds \emph{some} discrete loss $\ell^k$, and from Theorem~\ref{thm:eps-thick-calibrated} there is a link $\psi'$ such that $(L^k,\psi')$ is calibrated (and consistent) for $\ell^k$.
We therefore turn to deriving this discrete loss $\ell^k$.
